\documentclass{report}
\usepackage{fullpage}
\usepackage{url}
\begin{document}
\title{Progress Report for ManBearPig}
\author{spiros thanasoulas st19@illinois.edu}
\maketitle

\section*{Status}
As outlined in the project proposal the goal of this project is to document the current status and lay the foundation for providing
full text search capabilities to the mandoc system. 
Mandoc is a modern implementation of a manual page system \url{https://mandoc.bsd.lv} and part of the project is to identify its current
search capabilities (how the command apropos / man -k with a keyword works) and provide a design and parts of an implementation for
a full text search option on it.

\subsection*{Progress made so far}
I have analyzed the current way that apropos accepts search terms and passes them to the current database implementation.
Also i have designed an initial algorithm based on n-grams and a matching database format that could replace (or go side by side)
with the current database in order to provide some basic full text search capabilities. I have also written a scraper which 
can consume the mandoc language and extract the words that are to be fed to the database builder.


\subsection*{Remaining tasks}
Actually implement a database format that conforms to the apropos search api. 
Connect it to the apropos code and issue queries.
Write tests to make sure that the words that have been inputed have matching n-grams in the correct format in the 
binary database.
Write the report that describes the work done 

\subsection*{Challenges faced}
Being a large codebase it requires many hours of familiriazing with a part of it before you can start
making sense of how to interface with it.
Also being a pure C project , always requires treading carefully with things like memory management, IO etc
so the progress done is slower than it would be in a completely dynamic/scripting language working with a smaller framework.

\section*{members}
st19 / solo project 
\end{document}
